The Matter Neutrino Resonance around Thick, Relativisitic Disks

### 2016-03-23 paper.tex
Add JINA affiliation for Gail
Correct cosO.O -> O.O
Typos: remove a stray paren and switch a sign

The neutrino self-interaction term is
\begin{multline}
  \label{eq:H_nunu_full_form_1}
  \mathbb{H}_{\nu\nu}(s;\bm{\Omega}) \equiv
  \sqrt{2} G_F
  \int \frac{d^3p'}{(2\pi)^3}
  (1-\cos\Omega\cdot\Omega') \\
  \times
  \Big[\bbrho(s;\bm{p}')-\bar{\bbrho}(s;\bm{p}')\Big],
\end{multline}
where $\bm{p}'=\varepsilon' \bm{\Omega}'$ is an ambient neutrino
momentum. We can rewrite Eqn.~\ref{eq:H_nunu_full_form_1} using
Eqn.~\ref{eq:density_evolution} and
$d^3p'=\varepsilon'^2 d\varepsilon' d\Omega'$:
\begin{multline}
  \label{eq:H_nunu_full_form_2}
  \mathbb{H}_{\nu\nu}(s;\bm{\Omega}) =
  \frac{\sqrt{2} G_F}{(2\pi)^3}
  \oint d\Omega' \int d\varepsilon' \,\varepsilon'^2
  (1-\cos\Omega\cdot\Omega') \\
  \times \Big[
    \mathbb{S}(s;\varepsilon',\bm{\Omega}')
    \bbrho_0(s;\varepsilon',\bm{\Omega}')
    \mathbb{S}^\dagger(s;\varepsilon',\bm{\Omega}') \\
    -\bar{\mathbb{S}}^*(s;\varepsilon',\bm{\Omega}')
    \bar{\bbrho}_0(s;\varepsilon',\bm{\Omega}')
    \bar{\mathbb{S}}^T(s;\varepsilon',\bm{\Omega}')
    \Big].
\end{multline}
To compute this at a single point on our trajectory, $\bm{x}(s)$,
we need $\mathbb{S}(s)$ and $\bar{\mathbb{S}}(s)$ at all
energies $\varepsilon'$ and angles $\bm{\Omega}'$. In the formalism
we have developed thusfar, we can only get these by integrating
Eqn.~\ref{eq:S_evolution} along all trajectories in the family
sharing the point $\bm{x}(s)$; our problem is heavily recursive.

\subsection*{Single-angle approximation}
However, if we assume that every trajectory ($\bm{x}(s);\bm{\Omega}$)
yields the same flavor evolution, then we can make the substitution
\begin{equation}
 \mathbb{S}(s;\varepsilon',\bm{\Omega}') \rightarrow
 \mathbb{S}(s;\varepsilon',\bm{\Omega}), \nonumber
\end{equation}
and similarly for $\bar{\mathbb{S}}$, giving us
\begin{multline}
  \label{eq:H_nunu_single_angle}
  \mathbb{H}_{\nu\nu}(s;\bm{\Omega}) =
  \frac{\sqrt{2} G_F}{(2\pi)^3}
  \oint d\Omega' \int d\varepsilon' \,\varepsilon'^2
  (1-\cos\Omega\cdot\Omega') \\
  \times \Big[
    \mathbb{S}(s;\varepsilon',\bm{\Omega})
    \bbrho_0(s;\varepsilon',\bm{\Omega}')
    \mathbb{S}^\dagger(s;\varepsilon',\bm{\Omega}) \\
    -\bar{\mathbb{S}}^*(s;\varepsilon',\bm{\Omega})
    \bar{\bbrho}_0(s;\varepsilon',\bm{\Omega}')
    \bar{\mathbb{S}}^T(s;\varepsilon',\bm{\Omega})
    \Big].
\end{multline}

\subsection*{Self-interaction potentials}
Having applied the single-angle approximation, we can now factor
Eqn.~\ref{eq:H_nunu_single_angle} and perform the angle integration
before we compute $\mathbb{S}$ and $\bar{\mathbb{S}}$:
\begin{multline}
  \label{eq:H_nunu_factored}
  \mathbb{H}_{\nu\nu}(s;\bm{\Omega}) = \\
  \int d\varepsilon' \,
  \mathbb{S}(s;\varepsilon',\bm{\Omega})
  \bigg[
    \frac{1}{\Delta\varepsilon'}\mathbb{V}_0(s;\varepsilon',\bm{\Omega})
    \bigg]
  \mathbb{S}^\dagger(s;\varepsilon',\bm{\Omega}) \\
  -\int d\varepsilon' \,
  \bar{\mathbb{S}}^*(s;\varepsilon',\bm{\Omega})
  \bigg[
    \frac{1}{\Delta\varepsilon'} \bar{\mathbb{V}}_0(s;\varepsilon',\bm{\Omega})
    \bigg]
  \bar{\mathbb{S}}^T(s;\varepsilon',\bm{\Omega}).
\end{multline}
To simplify our equations we have defined
\begin{multline}
  \label{eq:nu_potential_matrix_1}
  \mathbb{V}_0(s;\varepsilon',\bm{\Omega}) \equiv \\
  \frac{\sqrt{2} G_F}{(2\pi)^3}\varepsilon'^2\Delta\varepsilon'
    \oint d\Omega' 
    (1-\cos\Omega\cdot\Omega')
    \bbrho_0(s;\varepsilon',\bm{\Omega}').
\end{multline}
And $\bar{\mathbb{V}}_0$ is defined similarly with respect to $\bar{\bbrho}_0$.
Note the new energy factor $\Delta\varepsilon'$ in our definition
in Eqn.~\ref{eq:nu_potential_matrix_1}:
this term gives $\mathbb{V}_0$ units of energy.
We will find it simplifies things when we sum $\mathbb{V}_0$ over energy bins.

@@

For completeness, the neutrino potentials in
Eqn.~\ref{eq:nu_potential_matrix_2} are defined
\begin{multline}
  \label{eq:nu_potential}
  V_{0,\nu_\alpha}(s;\varepsilon',\bm{\Omega}) \equiv \\
  \frac{\sqrt{2} G_F}{(2\pi)^3}\varepsilon'^2 \Delta\varepsilon'
  \oint d\Omega'
  (1-\cos\Omega\cdot\Omega')
  f_{0,\nu_\alpha}(s;\varepsilon',\bm{\Omega}'),
\end{multline}
This follows from Eqn.~\ref{eq:def_unoscillated_rho}.

@@

\subsection*{Oscillated self-interaction potential}
The Matter-Neutrino Resonance may occur along the test trajectory
where $\mathbb{H}_e$ and $\mathbb{H}_{\nu\nu}$ are the same
magnitude and differ by sign. If $\mathbb{S}$ and $\bar{\mathbb{S}}$ didn't
evolve from identity, the self-interaction Hamiltonian would take the form
\begin{equation}
  \mathbb{H}_{\nu\nu,0}(s;\bm{\Omega}) :=
  V_{\nu\nu,0}(s;\bm{\Omega})
  \left(
  \begin{matrix}
    1 & 0 & 0 \\
    0 & 0 & 0 \\
    0 & 0 & 0
  \end{matrix}
  \right),
\end{equation}
where
\begin{multline}
  V_{\nu\nu,0}(s;\bm{\Omega}) \equiv \\
  \sum\limits_i
  \big((V_{0,\nu_e}(s;\varepsilon'_i,\bm{\Omega})
  -V_{0,\bar{\nu}_e}(s;\varepsilon'_i,\bm{\Omega})\big).
\end{multline}
This is a nice quantity to compare to the matter potential
in Eqn.~\ref{eq:matter_potential}: an MNR resonance may occur at
$V_{\nu\nu,0}(s;\bm{\Omega})-V_e(s)\approx0$.

However, as flavor evolution does take place, $\mathbb{H}_{\nu\nu}$
develops a complex structure.
We define the oscillated self-interaction potential as
\begin{equation}
  V_{\nu\nu}(s;\bm{\Omega}) \equiv
  \big[\mathbb{H}_{\nu\nu}(s;\bm{\Omega})\big]_{ee}
  - \frac{1}{3}Tr\,\mathbb{H}_{\nu\nu}(s;\bm{\Omega}).
\end{equation}
As an MNR evolves, the flavor evolution seems to force this quantity to obey
$V_{\nu\nu}(s;\bm{\Omega})-V_e(s)\approx0$
(as long as possible).

>>>>

The neutrino self-interaction term is
\begin{multline}
  \label{eq:H_nunu_full_form_1}
  \mathbb{H}_{\nu\nu}(s;\bm{\Omega}) \equiv
  \sqrt{2} G_F
  \int \frac{d^3p'}{(2\pi)^3}
  (1-\Omega\cdot\Omega') \\
  \times
  \Big[\bbrho(s;\bm{p}')-\bar{\bbrho}(s;\bm{p}')\Big],
\end{multline}
where $\bm{p}'=\varepsilon' \bm{\Omega}'$ is an ambient neutrino
momentum. We can rewrite Eqn.~\ref{eq:H_nunu_full_form_1} using
Eqn.~\ref{eq:density_evolution} and
$d^3p'=\varepsilon'^2 d\varepsilon' d\Omega'$:
\begin{multline}
  \label{eq:H_nunu_full_form_2}
  \mathbb{H}_{\nu\nu}(s;\bm{\Omega}) =
  \frac{\sqrt{2} G_F}{(2\pi)^3}
  \oint d\Omega' \int d\varepsilon' \,\varepsilon'^2
  (1-\Omega\cdot\Omega') \\
  \times \Big[
    \mathbb{S}(s;\varepsilon',\bm{\Omega}')
    \bbrho_0(s;\varepsilon',\bm{\Omega}')
    \mathbb{S}^\dagger(s;\varepsilon',\bm{\Omega}') \\
    -\bar{\mathbb{S}}^*(s;\varepsilon',\bm{\Omega}')
    \bar{\bbrho}_0(s;\varepsilon',\bm{\Omega}')
    \bar{\mathbb{S}}^T(s;\varepsilon',\bm{\Omega}')
    \Big].
\end{multline}
To compute this at a single point on our trajectory, $\bm{x}(s)$,
we need $\mathbb{S}(s)$ and $\bar{\mathbb{S}}(s)$ at all
energies $\varepsilon'$ and angles $\bm{\Omega}'$. In the formalism
we have developed thusfar, we can only get these by integrating
Eqn.~\ref{eq:S_evolution} along all trajectories in the family
sharing the point $\bm{x}(s)$; our problem is heavily recursive.

\subsection*{Single-angle approximation}
However, if we assume that every trajectory ($\bm{x}(s);\bm{\Omega}$)
yields the same flavor evolution, then we can make the substitution
\begin{equation}
 \mathbb{S}(s;\varepsilon',\bm{\Omega}') \rightarrow
 \mathbb{S}(s;\varepsilon',\bm{\Omega}), \nonumber
\end{equation}
and similarly for $\bar{\mathbb{S}}$, giving us
\begin{multline}
  \label{eq:H_nunu_single_angle}
  \mathbb{H}_{\nu\nu}(s;\bm{\Omega}) =
  \frac{\sqrt{2} G_F}{(2\pi)^3}
  \oint d\Omega' \int d\varepsilon' \,\varepsilon'^2
  (1-\Omega\cdot\Omega') \\
  \times \Big[
    \mathbb{S}(s;\varepsilon',\bm{\Omega})
    \bbrho_0(s;\varepsilon',\bm{\Omega}')
    \mathbb{S}^\dagger(s;\varepsilon',\bm{\Omega}) \\
    -\bar{\mathbb{S}}^*(s;\varepsilon',\bm{\Omega})
    \bar{\bbrho}_0(s;\varepsilon',\bm{\Omega}')
    \bar{\mathbb{S}}^T(s;\varepsilon',\bm{\Omega})
    \Big].
\end{multline}

\subsection*{Self-interaction potentials}
Having applied the single-angle approximation, we can now factor
Eqn.~\ref{eq:H_nunu_single_angle} and perform the angle integration
before we compute $\mathbb{S}$ and $\bar{\mathbb{S}}$:
\begin{multline}
  \label{eq:H_nunu_factored}
  \mathbb{H}_{\nu\nu}(s;\bm{\Omega}) = \\
  \int d\varepsilon' \,
  \mathbb{S}(s;\varepsilon',\bm{\Omega})
  \bigg[
    \frac{1}{\Delta\varepsilon'}\mathbb{V}_0(s;\varepsilon',\bm{\Omega})
    \bigg]
  \mathbb{S}^\dagger(s;\varepsilon',\bm{\Omega}) \\
  -\int d\varepsilon' \,
  \bar{\mathbb{S}}^*(s;\varepsilon',\bm{\Omega})
  \bigg[
    \frac{1}{\Delta\varepsilon'} \bar{\mathbb{V}}_0(s;\varepsilon',\bm{\Omega})
    \bigg]
  \bar{\mathbb{S}}^T(s;\varepsilon',\bm{\Omega}).
\end{multline}
To simplify our equations we have defined
\begin{multline}
  \label{eq:nu_potential_matrix_1}
  \mathbb{V}_0(s;\varepsilon',\bm{\Omega}) \equiv \\
  \frac{\sqrt{2} G_F}{(2\pi)^3}\varepsilon'^2\Delta\varepsilon'
    \oint d\Omega' 
    (1-\Omega\cdot\Omega')
    \bbrho_0(s;\varepsilon',\bm{\Omega}').
\end{multline}
And $\bar{\mathbb{V}}_0$ is defined similarly with respect to $\bar{\bbrho}_0$.
Note the new energy factor $\Delta\varepsilon'$ in our definition
in Eqn.~\ref{eq:nu_potential_matrix_1}:
this term gives $\mathbb{V}_0$ units of energy.
We will find it simplifies things when we sum $\mathbb{V}_0$ over energy bins.

@@

For completeness, the neutrino potentials in
Eqn.~\ref{eq:nu_potential_matrix_2} are defined
\begin{multline}
  \label{eq:nu_potential}
  V_{0,\nu_\alpha}(s;\varepsilon',\bm{\Omega}) \equiv \\
  \frac{\sqrt{2} G_F}{(2\pi)^3}\varepsilon'^2 \Delta\varepsilon'
  \oint d\Omega'
  (1-\Omega\cdot\Omega')
  f_{0,\nu_\alpha}(s;\varepsilon',\bm{\Omega}'),
\end{multline}
This follows from Eqn.~\ref{eq:def_unoscillated_rho}.

@@

\subsection*{Oscillated self-interaction potential}
The Matter-Neutrino Resonance may occur along the test trajectory
where $\mathbb{H}_e$ and $\mathbb{H}_{\nu\nu}$ are the same
magnitude and differ by sign. If $\mathbb{S}$ and $\bar{\mathbb{S}}$ didn't
evolve from identity, the self-interaction Hamiltonian would take the form
\begin{equation}
  \mathbb{H}_{\nu\nu,0}(s;\bm{\Omega}) :=
  V_{\nu\nu,0}(s;\bm{\Omega})
  \left(
  \begin{matrix}
    1 & 0 & 0 \\
    0 & 0 & 0 \\
    0 & 0 & 0
  \end{matrix}
  \right),
\end{equation}
where
\begin{multline}
  V_{\nu\nu,0}(s;\bm{\Omega}) \equiv \\
  \sum\limits_i
  \big(V_{0,\nu_e}(s;\varepsilon'_i,\bm{\Omega})
  -V_{0,\bar{\nu}_e}(s;\varepsilon'_i,\bm{\Omega})\big).
\end{multline}
This is a nice quantity to compare to the matter potential
in Eqn.~\ref{eq:matter_potential}: an MNR resonance may occur at
$V_{\nu\nu,0}(s;\bm{\Omega})+V_e(s)\approx0$.

However, as flavor evolution does take place, $\mathbb{H}_{\nu\nu}$
develops a complex structure.
We define the oscillated self-interaction potential as
\begin{equation}
  V_{\nu\nu}(s;\bm{\Omega}) \equiv
  \big[\mathbb{H}_{\nu\nu}(s;\bm{\Omega})\big]_{ee}
  - \frac{1}{3}Tr\,\mathbb{H}_{\nu\nu}(s;\bm{\Omega}).
\end{equation}
As an MNR evolves, the flavor evolution seems to force this quantity to obey
$V_{\nu\nu}(s;\bm{\Omega})+V_e(s)\approx0$
(as long as possible).
